---
title: "EDS 212 Day 5 Activites"
author: "Halina Do-Linh"
date: "8/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Taking a look at `starwars`
```{r, include=FALSE}

# `include =` to show or hide everything
# use `echo =` to show or hide code (only outputs)
summary(starwars)

dim(starwars)

head(starwars)

tail(starwars)

names(starwars)

```

```{r, eval=FALSE}

#only use aes() when you are working with a variable. So if you wanted to change color by species use aes() because species is a variable.

# `eval =` to hide results, in this case, the graph.

# `starwars$name` shows us all the objects in the name column

ggplot(data = starwars, aes(x = height, y = mass)) +
  geom_point(color = "purple")

```


## Find the mean, variance, and SD of character height
```{r}

sw_height_mean <- mean(starwars$height, na.rm = TRUE)

# what are the measurements? cm

# We get a NA value as the return. Why? Because there are missing values. And the default in the `mean()` is to not remove the NA values. 

# To remove NA values you use `na.rm = TRUE`

# Using similar naming, find:
# Variance of character height (var())
# Standard deviation of character height (sd())

sw_height_variance <- var(starwars$height, na.rm = TRUE)

# what are the measurements? cm^2

sw_height_sd <- sd(starwars$height, na.rm = TRUE)

# what are the measurements? cm

```

Mean character height is 174.4 cm with a standard deviation of 34.8 cm.

Why is this a bad sentence? Because these are hard numbers and won't be able to update if mistakes are found or if you need to add/update data or to update a version. Also no one else could go back and see what code you wrote to get these numbers. 

Solution: R in line using back ticks

Mean character height is `r round(sw_height_mean, 1)` cm, with a standard deviation of `r round(sw_height_sd, 1)` cm.

Protip: only round values in the reporting stage so you don't crop your analysis early on. 

## Logicals Expressions
```{r}

# logical expressions ask questions and return TRUE or FALSE. 

5 < 10

"cat" > "armadillo"

"cat" > "elephant"

"dog" == "dog"

```


```{r}
10 <= 11

5 >= 4

3 <= 1

```

```{r}
tarantulas <- c(1, 3, 7, 10, 15)

tarantulas > 5

```


## Logical Operators 

- `!`
- `&`
- `|`

```{r}

3 < 1 | 8 > 2

# Returned true because one is true

6 < 10 & 5 < 1

# Returned false only one is true

7 != 11

# This is asking "Is 7 not equal to 11?" Returned true.

4 != 4 | 2 < 5 | 6 > 10

# This is a Union statement. So as long as one is true the whole statement is  true.

0 == FALSE

1 == TRUE

# R automatically reads 0 as false and 1 as true.

```

